Analyse and data mining of a chromatograph during drilling operations

### 2018-11-30 papers/chromatography (for ethics course)/paper.tex
Chromatography paper: introduction and first part with hardware added

\title{Command, analyse and data mining of a chromatograph during drilling operations
}

@@

@@

The IEEEtran class file is used to format your paper and style the text. All margins, 
column widths, line spaces, and text fonts are prescribed; please do not 
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement 
and others are deliberate, using specifications that anticipate your paper 
as one part of the entire proceedings, and not as an independent document. 
Please do not revise any of the current designations.

>>>>

\title{Analyse and data mining of a chromatograph during drilling operations}

@@

\subsection{Chromatography concepts}

    Gas chromatography means to separate components such that they can be analyzed independently. To achive this a
    separation column is used. A probe of gas, taken from drilling installation, is pushed through the column which
    means that all components enter into it at the same time. The column is filled with carefully choosen particles,
    whose type and size are competitive commercial advantages (and because this is not the object of the current paper,
    their details will not be disclosed). The probe is pushed through the column using a bearing gas, usually hydrogen.
    The probe has to make its way through the column's particles and because of the friction between them different
    components will have different speeds with which they advance. Thus, the components are reaching the end of the 
    column at different times. The time which is needed for a component to traverse the column is called retention time.
    Figure 1 contains a representation of a column and how two components, A and B, enter into the column at the same
    time but they exit at different times.
    During the usual drilling operations geologist engineers are interested in certain components so that they can analyze
    and predict the behavior of the gas or petrol reservoir. Those components are: methane (C1), ethane (C2), propane (C3),
    n-butane (nC4), isobutane (iC4), neopentane (nC5), isopentane (iC5). The sum of all those components yields a number
    called theoretical total gas. The probe is also analysed as a whole to find out the total gas. The ratio of those two
    numbers is of paramount importance.
    Among the details which can be find out using the analyses of those components are the quality of the gas or petrol
    which is in the reservoir, the proximity of the reservoir or certain predictors important for drilling engineers, for
    example, the danger of the entire reservoir to erupt.

    \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{column.png}
        \caption{Gas separation column}
    \end{figure}

\subsection{Flame Ionisation Detector}

    The components are slowly (or fast, depending on the performance of the chromatograph) pouring out of the column
    into a detector called Flame Ionisation Detector or FID. Its role is to transform over the analyses time the quantity
    of each component into voltage. This voltage is read by ADC boards and transmitted to a computer for further calculations.
    Figure 2 shows the schematics of such a FID. Item 1 represents the separation column, item 2 represents suplimentary
    gas supply, 3 represents hydrogen supply, 4 represents the lighter, 5 represents the collector electrode, 6 is the
    polarizing electrode, 7 is the burner and 8 is the air supply.

    \begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{fid.png}
        \caption{Flame Ionisation Detector}
    \end{figure}

@@

Following description will be brief one and it focuses only on what is important for this paper. The construction of a 
chromatograph implies certain components to be glued together.
